Supporting Information\\ \textit{Cross Validation Approaches for Penalized Cox Regression}

### 2020-03-30 main.tex
baseline info in supplementary

\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.  In doing so, we calculate the risk for each member of the test set relative to other members of the test set.  One drawback to this approach is that it becomes unstable when the size of the the test set is small.  In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.

@@

\bnote{Simulations with features that has more complicated correlation structures were conducted as well. Similar patterns were observed as in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim}. Results of these simulations are provided in the supporting document.}

>>>>

\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.  In doing so, we calculate the risk for each member of the test set relative to other members of the test set.  One drawback to this approach is that it becomes unstable when the size of the test set is small.  In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.

@@

\bnote{More simulations were conducted to study scenarios where features have more complicated correlation structures. Similar patterns were observed as in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim}. Results of these simulations are provided in the supplementary material.}

### 2020-03-30 supplementary.tex
baseline info in supplementary

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

@@

	\subsubsection{Nelson-Aalen Estimator}
	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
	The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times.

@@

	\subsubsection{Kalbfleisch and Prentice's Approach}					
	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates. Assuming only a single failure occurs at $t_i$, based on the fitted values of $\hat{\bbeta}$, we can first compute linear predictors $\hat{\eta}_i$ at each observed failure time. Based on $\hat{\eta}_i$, we can compute a discrete hazard component at each observed failure time):
\begin{equation}
1 - \hat{\alpha}_{i} = 1 - \left\{ 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}_{j}} \right\}^{exp(-\hat{\eta}_{i})},
\end{equation}
and aggregate them together into the cumulative hazard function:
\begin{equation} 
\hat{\Lambda}_0(t) = \sum(1 - \hat{\alpha}_i)\mathds{1}(t_{i} \leq t),
\end{equation}

>>>>

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M}^{cv}_{i} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

@@

	\subsubsection{Nelson-Aalen Estimator}
	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
Under the assumption that there is no ties among failure times, the numerator in the summation is either 0 or 1. The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times. When we use the Nelson-Aalen Estimator for the cross-validated deviance residuals, we follow the below algorithm:
\begin{algorithm}
\SetAlgoLined
Use all observations' time and status to compute $\hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
\For{each regularization parameter $\lambda$}{
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{Compute Cross-validated Deviance Residuals Based on Nelson-Aalen Estimator}
\end{algorithm}

@@

In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard. Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
\begin{equation} 
\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
\end{equation}
Covariates of the fitted model are incorporated in the denominator of the summation. When we use Breslow's estimator for the cross-validated deviance residuals, we follow the below algorithm:
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 Compute $\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$\;
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{Compute Cross-validated Deviance Residuals Based on Breslow's Approach}
\end{algorithm}

\subsubsection{Kalbfleisch and Prentice's Approach}

	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates. Assuming only a single failure occurs at $t_i$, based on the fitted values of $\hat{\bbeta}$, we can first compute linear predictors $\hat{\eta}_i$ at each observed failure time. Based on $\hat{\eta}_i$, we can compute a discrete hazard component at each observed failure time):
\begin{equation}
1 - \hat{\alpha}_{i} = 1 - \left( 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right)^{\exp(-\hat{\eta}_{i})},
\end{equation}
and aggregate them together into the cumulative hazard function:
\begin{equation} 
\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}(1 - \hat{\alpha}_i),
\end{equation}
When we use the Kalbfleisch and Prentice's Estimator for the cross-validated deviance residuals, we follow the below algorithm:
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 Compute $\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{Compute Cross-validated Deviance Residuals Based on Kalbfleisch and Prentice's Estimator}
\end{algorithm}
